---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

```{r}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

Loading packages

```{r, message = FALSE}
library(ggplot2)
library(dplyr)
library(lattice)
```

## Loading and preprocessing the data

```{r total numerber}
activity <- read.csv("./activity.csv")
```


## What is mean total number of steps taken per day?

```{r histgram, fig.width = 10, fig.height=5}
activity <- activity[which(!is.na(activity$steps)), ]
sum_steps <- aggregate(steps~date, activity, sum, na.rm = TRUE)
g <- ggplot(sum_steps, aes(x = date, y = steps))
g + geom_histogram(stat='identity') + theme(axis.text.x = element_text(angle=90, hjust=1, vjust=0.5)) + labs(title = "The total number of steps taken each day") +  theme(plot.title = element_text(hjust = 0.5, size = 15))
```

## Mean and median number of steps taken each day

```{r mean and median number of steps}
group_activity <- group_by(activity, date)
mean_steps <- data.frame(summarize(group_activity, mean(steps)))
median_steps <- data.frame(summarize(group_activity, median(steps)))
mean_median_steps <- merge(mean_steps, median_steps, by.x = "date", by.y = "date", all.x = TRUE)
mean_median_steps
```

## What is the average daily activity pattern?
### Time series plot of the average number of steps all the days

```{r average number of steps, fig.width=10}
group_activity <- group_by(activity, interval)
mean_interval_steps <- data.frame(summarize(group_activity, mean(steps)))
mean_interval_steps <- rename(mean_interval_steps, steps = "mean.steps.")
g <- ggplot(mean_interval_steps, aes(x = interval, y = steps))
g + geom_line(size = 1) + labs(title = "The average number of steps per 5 minutes") + theme(plot.title = element_text(hjust = 0.5, size = 15))
```

## The 5-minute interval that, on average, contains the maximum number of steps

```{r maximum number of steps}
max_steps_interval <- filter(mean_interval_steps, steps == max(mean_interval_steps$steps))$interval
```

## Imputing missing values

```{r imputing missing data}
activity <- read.csv("./activity.csv")
missing_value <- activity[is.na(activity$steps), ]
integer_interval_steps <- round(mean_interval_steps)
Imputing_missing_value <- merge(missing_value, integer_interval_steps, by.x = "interval", by.y = "interval", all.x = TRUE)
missing_value$id <- 1:nrow(missing_value)
Imputing_missing_value <- arrange(merge(missing_value, integer_interval_steps, by.x = "interval", by.y = "interval", all.x = TRUE), id)
Imputing_missing_value <- Imputing_missing_value[ ,c(-2,-4)]
noNa_activity <- activity
noNa_activity[is.na(noNa_activity$steps), ]$steps <- Imputing_missing_value$steps.y
noNa_activity <- noNa_activity
```

## Histogram of the total number of steps taken each day after missing values are imputed

```{r histgram for total numer os steps after missing values are imputed, fig.width=10}
noNa_sum_steps <- aggregate(steps~date, noNa_activity, sum)
g <- ggplot(noNa_sum_steps, aes(x = date, y = steps))
g + geom_histogram(stat='identity') + theme(axis.text.x = element_text(angle=90, hjust=1, vjust=0.5)) + labs(title = "The total number of steps taken each day after imputing missing value") +  theme(plot.title = element_text(hjust = 0.5, size = 15))
noNA_group_activity <- group_by(noNa_activity, date)
noNA_mean_steps <- data.frame(summarize(noNA_group_activity, mean(steps)))
noNA_median_steps <- data.frame(summarize(noNA_group_activity, median(steps)))
```

## Are there differences in activity patterns between weekdays and weekends?
1. Compared to weekends, people get up earlier on workdays
2. On weekends, the number of steps from morning (after commute) to afternoon is more than those on weekdays

temp <- merge(noNA_mean_steps, mean_steps, by.x = "date", by.y = "date", all.x = TRUE)
temp <- merge(temp, noNA_median_steps, by.x = "date", by.y = "date", all.x = TRUE)
temp <- merge(temp, median_steps, by.x = "date", by.y = "date", all.x = TRUE)
Imputation_contrast <- rename(temp, meansteps_afterImputation = mean.steps..x, meansteps_beforeImputation = mean.steps..y, mediansteps_afterImputation = median.steps..x, mediansteps_beforeImputation= median.steps..y)
Sys.setlocale("LC_TIME", "English")
noNa_activity$date <- as.Date(noNa_activity$date)
noNa_activity$label <- weekdays(noNa_activity$date, abbreviate = TRUE)
noNa_activity[which(grepl("S(at|un)", noNa_activity$label)), ]$label <- "weekend"
noNa_activity[(noNa_activity$label != "weekend"), ]$label <- "weekday"
noNa_activity$label <- as.factor(noNa_activity$label)
plot <- aggregate(noNa_activity[,1], by = list(interval = noNa_activity$interval, label = noNa_activity$label), FUN = mean)
plot <- rename(plot, steps = x)
xyplot(steps ~ interval| label, data = plot, layout = c(1, 2), type = "l", main = "The average number of steps taken per 5-minute interval across weekdays and weekends")
```
